---
title       : Estimating an iris petal length
subtitle    : Coursera project for Developing Data Products
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## The iris petal length predictor

1. Unique web application
2. Objective: Estimate the petal length of an iris from its species and its petal width
3. Highly interactive
4. Powerful explanatory graphics
5. Public and free !


https://patrocle-troy.shinyapps.io/project/

--- .class #id 

## The methodology

1. R programming language
2. Data set: Edgar Anderson's Iris Data (R-base dataset)
3. Prediction algorithm: linear regression 
  1. Length against width
  2. One regression per species

---

## The code
* Function: `predictLength` = linear regression for each species
* Input: `inSpecies` = species (as an integer), `inWidth` = petal width (in cm)
* Output: `dpredict` = iris dataframe with the estimated petal length
``` {r}
predictLength<-function(inSpecies,inWidth)
{ inSpecFact<-switch(inSpecies,'1'='setosa','2'='versicolor','3'='virginica');
  df = iris[,c('Petal.Length','Petal.Width','Species')]; colnames(df)<-c('length','width','species');
  lmselect<-lm(length~width,data=df[df$species==inSpecFact,]); 
  outLength<-predict(lmselect,data.frame(width=inWidth));
  dpredict<-data.frame(width=inWidth,length=outLength,species=inSpecFact); }
```
Example: `inSpecies=2` (versicolor); `inWidth=2.1`; output = length 
```{r}
print(predictLength(2,2.1)$length)
```

--- 
## The plot

```{r echo = FALSE}
  library(ggplot2)
  inSpecFact = 'versicolor'
  inWidth = 2.1
  df = iris[,c('Petal.Length','Petal.Width','Species')];
  colnames(df)<-c('length','width','species')
  lmselect<-lm(length~width,data=df[df$species==inSpecFact,])
  outLength<-predict(lmselect,data.frame(width=inWidth))
  dpredict <- data.frame(width=inWidth,length=outLength,species=inSpecFact)
  dplot<-rbind(dpredict,df)
  p<-qplot(width,length,data=dplot, col=species, main='Iris petals')+geom_point(aes(colour=dpredict$species), size=5,data = dpredict)
  print(p)
```




